---
title: "FinalProj"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(dplyr.summarise.inform = FALSE)

library(ggpubr)
library(betareg)
```

WHO Data: https://apps.who.int/gho/data/node.main.A1366?lang=en
Worldbank Data: https://data.worldbank.org/indicator/SH.MLR.INCD.P3
Prevention Data: https://data.unicef.org/resources/data_explorer/unicef_f/?ag=UNICEF&df=GLOBAL_DATAFLOW&ver=1.0&dq=AGO+BEN+BWA+BFA+BDI+CPV+CMR+CAF+TCD+COM+COG+COD+DJI+GNQ+ERI+ETH+GAB+GMB+GHA+KEN+LSO+LBR+MDG+MWI+MLI+MUS+MRT+MOZ+NAM+NER+NGA+RWA+STP+SEN+SYC+SLE+SOM+ZAF+SSD+SDN+TGO+UGA+TZA+ZMB+ZWE.MNCH_ITN+MNCH_ITN2+MNCH_ITNOWN+MNCH_ITNPREG+MNCH_MLRACT..&startPeriod=2016&endPeriod=2022


Load Country Case data.
Remove unnecessary columns and year columns with no data. 
Replace year column names with just the year.
```{r}
country.case.dat<-read.csv("API_SH.MLR.INCD.P3_DS2_en_csv_v2_4029728.csv")
country.case.dat<-country.case.dat[, -c(5:44, 66:67) ]
country.case.dat<- country.case.dat %>%
  select(-Indicator.Code)
names(country.case.dat) <- gsub(x = names(country.case.dat), pattern = "X", replacement = "")
country.case.dat<-country.case.dat %>%
  select(-Indicator.Name)
```

Load indicator data.
Change naming convention of some year columns.
Make new country code column to link to country case data, change geographic area to country name.
Change rows to total b/w male and female.
Remove unused indicators from indicator column.
Select only used columns
Left Join Data, pivot year column 
```{r}
indicator.dat<-read.csv("Prevention.csv")
indicator.dat$Country.Code=substr(indicator.dat$REF_AREA.Geographic.area, 1, 3)
indicator.dat$Country.Name=substring(indicator.dat$REF_AREA.Geographic.area, 6)
indicator.dat$TIME_PERIOD.Time.period[indicator.dat$TIME_PERIOD.Time.period=="2017-06"]<-"2017"

indicator.dat<-indicator.dat[indicator.dat$SEX.Sex=="_T: Total",]
indicator.dat<- indicator.dat[
  str_detect(indicator.dat$INDICATOR.Indicator,"MNCH_ITNOWN") |
  str_detect(indicator.dat$INDICATOR.Indicator,"MNCH_ITN2"),]
indicator.dat<-indicator.dat %>%
  select(c(Country.Name, OBS_VALUE.Observation.Value, INDICATOR.Indicator, Country.Code, TIME_PERIOD.Time.period))

full.dat<-left_join(indicator.dat, country.case.dat, by="Country.Code")

full.dat<- full.dat %>% 
  pivot_longer(c(7:27), names_to = "Year", values_to = "Cases")
```

left join columns
get means of malaria cases by year
QQplot
test data
Compare % with ITN and/or IRS with malaria cases that year and the following year
```{r}
yearly_means <- full.dat %>%
  group_by(Year) %>%
  summarize(mean=mean(Cases))

country_means <- full.dat %>%
  group_by(Country.Code) %>%
  summarize(mean=mean(Cases))

ggplot(yearly_means, aes(x=Year, y=mean)) +
  geom_line() +
  geom_point()
```


Here we're grouping by before the date of measurement of either 
 insecticide treated net (ITN) and/or indoor residual spraying (IRS), and getting the case mean,
 and the year of measurement + the year after and getting the mean, then combining. 
```{r}
pre.itn.dat <- full.dat %>%
  filter(str_detect(INDICATOR.Indicator,"MNCH_ITNOWN")) %>%
  filter(Year<TIME_PERIOD.Time.period) %>%
  group_by(Country.Code) %>%
  mutate(Pre.Measure.Case.Mean=mean(Cases)) %>%
  mutate(Pre.Measure.Case.Mean.Prop = Pre.Measure.Case.Mean/1000) %>%
  mutate(Percent.Households.With.ITN = OBS_VALUE.Observation.Value/100) %>%
  select(Country.Name, Pre.Measure.Case.Mean, Pre.Measure.Case.Mean.Prop, Percent.Households.With.ITN) %>%
  distinct()

post.itn.dat <- full.dat %>%
  filter(str_detect(INDICATOR.Indicator,"MNCH_ITNOWN") &
    Year %in% c(TIME_PERIOD.Time.period, as.numeric(TIME_PERIOD.Time.period)+1)) %>%
  group_by(Country.Code) %>%
  mutate(Post.Measure.Case.Mean=mean(Cases)) %>%
  mutate(Post.Measure.Case.Mean.Prop = Post.Measure.Case.Mean/1000) %>%
  mutate(Percent.Households.With.ITN = OBS_VALUE.Observation.Value/100) %>%
  select(Country.Code, Post.Measure.Case.Mean, Post.Measure.Case.Mean.Prop, Percent.Households.With.ITN) %>%
  distinct()

full.itn.dat<-merge(pre.itn.dat, post.itn.dat)

ggqqplot(full.itn.dat$Pre.Measure.Case.Mean-full.itn.dat$Post.Measure.Case.Mean)
```

```{r}


pre.itn.irs.dat <- full.dat %>%
  filter(str_detect(INDICATOR.Indicator,"MNCH_ITN2")) %>%
  filter(Year==TIME_PERIOD.Time.period) %>%
  group_by(Country.Code) %>%
  mutate(Pre.Measure.Case.Mean=mean(Cases)) %>%
  mutate(Pre.Measure.Case.Mean.Prop = Pre.Measure.Case.Mean/1000) %>%
  mutate(Percent.Households.With.ITN.Or.IRS = OBS_VALUE.Observation.Value/100) %>%
  select(Country.Code, Pre.Measure.Case.Mean, Pre.Measure.Case.Mean.Prop, Percent.Households.With.ITN.Or.IRS) %>%
  distinct()

post.itn.irs.dat <- full.dat %>%
  filter(str_detect(INDICATOR.Indicator,"MNCH_ITN2") &
    Year %in% c(TIME_PERIOD.Time.period, as.numeric(TIME_PERIOD.Time.period)+1)) %>%
  group_by(Country.Code) %>%
  mutate(Post.Measure.Case.Mean=mean(Cases)) %>%
  mutate(Post.Measure.Case.Mean.Prop = Post.Measure.Case.Mean/1000) %>%
  mutate(Percent.Households.With.ITN.Or.IRS = OBS_VALUE.Observation.Value/100) %>%
  select(Country.Code, Post.Measure.Case.Mean, Post.Measure.Case.Mean.Prop, Percent.Households.With.ITN.Or.IRS) %>%
  distinct()

full.itn.irs.dat<-merge(pre.itn.irs.dat, post.itn.irs.dat)
ggqqplot(full.itn.irs.dat$Pre.Measure.Case.Mean-full.itn.irs.dat$Post.Measure.Case.Mean)
```
```{r}
ggplot(full.itn.irs.dat, aes(x=Percent.Households.With.ITN.Or.IRS, y=Post.Measure.Case.Mean)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE)
ggplot(full.itn.dat, aes(x=Percent.Households.With.ITN, y=Post.Measure.Case.Mean)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
betareg(Post.Measure.Case.Mean.Prop ~ Percent.Households.With.ITN.Or.IRS, data=full.itn.irs.dat)
```
